Fixing Vulnerabilities Hinders Maintainability

### 2020-01-31 paper/icpc20.tex
new version

As mentioned before, this study evaluates $1037$ security and baseline commits
from $267$ distinct open-source projects. The following section presents the
results obtained for each research question.

@@

Maintainability is being calculated as the mean of the guidelines computed
by BCH. Figure~\ref{fig:guidelines} shows the impact of security patches on each 
guideline individually. For projects with large 
codebases, the results retrieved for \emph{Keep Your Codebase Small} were not 
viable due to limitations of BCH. For this guideline, the tool sets a limit of 20 Person-years. However, a large amount of projects of our database surpass this limit. Thus, we decided to not consider this guideline 
in our analysis. For each type of
guideline, three types of impact of patches in the software maintainability
are presented: \emph{negative}, maintainability decreases (\emph{red});
\emph{none}, maintainability remains the same (\emph{yellow}); and,
\emph{positive}, maintainability increases (\emph{green}). Next to each
type of guideline, it is shown the mean ($\overline{x}$) and median (M)
of the maintainability difference and the p-value resulting from the
Paired Wilcoxon signed-rank test.
Regarding security commits results, we observe that maintainability decreases in
$41.76\%$ ($433$), remains the same in $19.67\%$ ($204$) and increases in $38.57\%$
($400$) of the cases after security patches are applied. The resulting
p-value of the Paired Wilcoxon signed-rank test is $4.78$x$10^{-2}$.
Since the p-value is below the significance level of $0.05$, we argue that
\emph{security patches have a negative impact in the maintainability of
open-source software}.
Regarding the impact of security patches for each guideline, we observe
that overall patches have a significant positive impact in $5$ guidelines: 
\emph{Write Clean Code} (p-value = $4.9$x$10^{-54}$), \emph{Keep Arschitecture 
Components Balanced} (p-value = $5.26$x$10^{-71}$), \emph{Couple Architecture 
Components Loosely} (p-value = $2.43$x$10^{-07}$), \emph{Keep Unit Interfaces 
Small} (p-value = $2.25$x$10^{-13}$), \emph{ Write Code Once} (p-value = 
$7.9$x$10^{-35}$). However, for each of these guidelines we also observe a substantial number of cases where the maintainability suffers a negative impact $-$ from $10\%$ to $35\%$ $-$ which highlights the need of better practices/techniques/tools to produce software. \emph{Separate Concerns in Module} 
guideline also follows the same results but without any statistical 
significance since the p-value is higher than $0.05$. According with the 
results, \emph{Write Simple Units of Code} (p-value=$0.015$ and $38.2\%$/$396$ 
of negative cases) and \emph{Write Short Units of Code} (p-value=$0.029$ and 
$38.3\%$/$397$ of negative cases) are the guidelines which suffered more 
impact from security patches.

@@

For the vulnerabilities identified with CVE(s), we scraped NVD to retrieve 
their severity levels. In total, we obtained severity scores for $574$ 
vulnerabilities: $121$ \emph{HIGH}, $423$ \emph{MEDIUM} and $30$ \emph{LOW}. 
Figure~\ref{fig:severity} presents the impact
of security patches per severity level on the maintainability of open-source 
software. Although with no statistical significance, we observe that 
vulnerabilities with higher severity levels (\emph{HIGH} and \emph{MEDIUM}) have more negative impact than the \emph{LOW} severity level on the
maintainability of software.

@@

We have also evaluated the impact on the maintainability per programming language
of open-source software, Figure~\ref{fig:lang_main}. Although
the dataset contains security patches for more languages, we only
present the results for \emph{Python}, \emph{C/C++}, \emph{Ruby}, \emph{PHP}, 
\emph{Objective-C/C++}, and \emph{Java}
because of the Paired Wilcoxon signed-rank test requirement: input sample has to have more 
than $20$ samples. The other languages were
integrated into \emph{Others}. The results of the Paired Wilcoxon signed-rank
test by language indicate that there is statistical evidence to support that the
maintainability of open-source software decreases for security patches for \emph{C/C++} 
(p-value = $2.5$x$10^{-5}$). 
% Following the \textit{American Statistical Association}
% recommendations~\cite{doi:10.1080/00031305.2016.1154108}, we consider the results for
% Ruby as statistically significant because its p-value is on the borderline.
We could also conclude that security 
patches in \emph{PHP} (p-value = $0.072$) also have a negative impact
on the maintainability of software, if we considered a significance level of $0.1$. 
Another programming language which impacts negatively the maintainability of software
is \emph{Ruby} (30 in 71 vulnerabilities, $42.3\%$).

@@

Software with \emph{Resource Management Errors (CWE-399)} flaws are the most affected ones by security patches: $54.5\%$ ($12$) of the patches had a negative impact
in the maintainability while $36.4\%$ ($8$) yielded a positive impact, and only
$9.1\%$ ($2$) did not suffer any change. The results also show that fixing
\emph{Information Exposure (CWE-200)} vulnerabilities affects the maintainability of $51.2\%$ ($20$) of the cases
and leaves only $2.6\%$ ($1$) of the cases unaffected. In $46.2\%$ ($18$) of the
cases, maintainability increased. 
Patching \emph{Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (CWE-22)} vulnerabilities has a negative
impact of $48.6\%$ ($18$) in the maintainability of open-source software while
$46.0\%$ ($17$) of the cases exhibit maintainability improvement and $5.4\%$
($2$) show that maintainability does not suffer any change. The
\emph{Miscellaneous} patches pattern integrates several types of security
flaws. These type of patches harm maintainability in $41.3\%$ ($673$) of the
cases. In $22.7\%$ ($153$), there is no impact and for the
other $36.0\%$ ($242$) maintainability increases.
It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting (CWE-79)} vulnerabilities. \emph{Cross-Site Scripting (CWE-79)}
does not suffer any decrease in the maintainability of $38.7\%$ ($24$) of cases
and $35.5\%$ ($22$) have a negative effect on it. Only $25.8\%$ ($16$) of the
cases improve the software maintainability.

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.43\textwidth]{figures/maintainability_cwe.pdf}
 	\caption{Maintainability difference by pattern after security patches.}
	\label{fig:pat}	
	\vspace{-0.5cm}
\end{figure}

>>>>

This study evaluates $1037$ security and baseline commits
from $267$ distinct open-source projects. The following section presents the
results obtained for each research question.

@@

Maintainability is being calculated as the mean of the guidelines computed
by BCH. Figure~\ref{fig:guidelines} shows the impact of security patches on each 
guideline individually. For projects with large 
codebases, the results retrieved for \emph{Keep Your Codebase Small} were not 
viable due to limitations of BCH. For this guideline, the tool sets a limit of 20 Person-years. However, a large amount of projects of our database surpass this limit. Thus, we decided to not consider this guideline 
in our analysis. For each type of
guideline, three types of impact of patches in the software maintainability
are presented: \emph{negative}, maintainability decreases (\emph{red});
\emph{none}, maintainability remains the same (\emph{yellow}); and,
\emph{positive}, maintainability increases (\emph{green}). Next to each
type of guideline, it is shown the mean ($\overline{x}$) and median (M)
of the maintainability difference and the p-value resulting from the
Paired Wilcoxon signed-rank test.

Regarding security commits results, we observe that maintainability decreases in
$41.76\%$ ($433$), remains the same in $19.67\%$ ($204$) and increases in $38.57\%$
($400$) of the cases after security patches are applied. The resulting
p-value of the Paired Wilcoxon signed-rank test is $4.78$x$10^{-2}$.
Since the p-value is below the significance level of $0.05$, we argue that
\emph{security patches have a negative impact in the maintainability of
open-source software}.
Regarding the impact of security patches for each guideline, we observe
that overall patches have a significant positive impact in $5$ guidelines: 
\emph{Write Clean Code} (p-value = $4.9$x$10^{-54}$), \emph{Keep Architecture 
Components Balanced} (p-value = $5.26$x$10^{-71}$), \emph{Couple Architecture 
Components Loosely} (p-value = $2.43$x$10^{-07}$), \emph{Keep Unit Interfaces 
Small} (p-value = $2.25$x$10^{-13}$), \emph{ Write Code Once} (p-value = 
$7.9$x$10^{-35}$). However, for each of these guidelines we also observe a substantial number of cases where the maintainability suffers a negative impact $-$ from $10\%$ to $35\%$ $-$ which highlights the need of better practices/techniques/tools to produce software. \emph{Separate Concerns in Module} 
guideline also follows the same results but without any statistical 
significance since the p-value is higher than $0.05$. According to the 
results, \emph{Write Simple Units of Code} (p-value=$0.015$ and $38.2\%$/$396$ 
of negative cases) and \emph{Write Short Units of Code} (p-value=$0.029$ and 
$38.3\%$/$397$ of negative cases) are the most affected affected guidelines.

@@

For the vulnerabilities identified with CVE(s), we scraped the National Vulnerability Database (NVD) Website to retrieve 
their severity levels. In total, we obtained severity scores for $574$ 
vulnerabilities: $121$ \emph{HIGH}, $423$ \emph{MEDIUM} and $30$ \emph{LOW}. 
Figure~\ref{fig:severity} presents the impact
of security patches per severity level on the maintainability of open-source 
software. Although with no statistical significance, we observe that 
vulnerabilities with higher severity levels (\emph{HIGH} and \emph{MEDIUM}) have more negative impact than the \emph{LOW} severity level on the
maintainability of software.

@@

We have also evaluated the impact on the maintainability per programming language
of open-source software, Figure~\ref{fig:lang_main}. We restrict this analysis to programming languages with at least $20$ data points, as this is a requirement for our hypothesis tests. Thus, we compare the results for \emph{Python}, \emph{C/C++}, \emph{Ruby}, \emph{PHP}, 
\emph{Objective-C/C++}, and \emph{Java}. The other languages were
integrated into \emph{Others}. The results of the Paired Wilcoxon signed-rank
test by language indicate that there is statistical evidence to support that the
maintainability of open-source software decreases for security patches for \emph{C/C++} 
(p-value = $2.5$x$10^{-5}$). 
% Following the \textit{American Statistical Association}
% recommendations~\cite{doi:10.1080/00031305.2016.1154108}, we consider the results for
% Ruby as statistically significant because its p-value is on the borderline.
\emph{Ruby} and \emph{PHP} languages also impact negatively the maintainability of software. Despite not having statistical significance, there are a considerable number of \emph{Ruby} and \emph{PHP} patches hindering maintainability that should not be disregarded

@@

Software with \emph{Resource Management Errors (CWE-399)} flaws are the most affected ones by security patches: $54.5\%$ ($12$) of the patches had a negative impact
in the maintainability while $36.4\%$ ($8$) yielded a positive impact, and only
$9.1\%$ ($2$) did not suffer any change. The results also show that fixing
\emph{Information Exposure (CWE-200)} vulnerabilities affects the maintainability of $51.2\%$ ($20$) of the cases
and leaves only $2.6\%$ ($1$) of the cases unaffected. In $46.2\%$ ($18$) of the
cases, maintainability increased. 
Patching \emph{Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (CWE-22)} vulnerabilities has a negative
impact of $48.6\%$ ($18$) in the maintainability of open-source software while
$46.0\%$ ($17$) of the cases exhibit maintainability improvement and $5.4\%$
($2$) show that maintainability does not suffer any change. The
\emph{Miscellaneous} patches pattern integrates several types of security
flaws. These type of patches harm maintainability in $41.3\%$ ($673$) of the
cases. In $22.7\%$ ($153$), there is no impact and for the
other $36.0\%$ ($242$) maintainability increases.
It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting (CWE-79)} vulnerabilities. \emph{Cross-Site Scripting (CWE-79)}
does not suffer any decrease in the maintainability of $38.7\%$ ($24$) of the cases
and $35.5\%$ ($22$) have a negative effect on it. Only $25.8\%$ ($16$) of the
cases improve the software maintainability.

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.4\textwidth]{figures/maintainability_cwe.pdf}
 	\caption{Maintainability difference by pattern after security patches.}
	\label{fig:pat}	
	\vspace{-0.5cm}
\end{figure}
